---
title: "JAGS analysis of population parameters"
---

Load packages
```{r}
library(tidyverse)
library(R2jags)
library("mcmcplots")
```

##Load data: lifetable data
```{r}
pheno_data0 <- read_csv("../data/final_phenotype_data.csv",
                       col_names = TRUE,
                       col_types = list(host = col_factor(NULL),
                                        pathogen = col_factor(levels = c("CTRL", "C1", "C14", "C20")),
                                        temp_treat = col_factor(levels = c("20-20", "25-20", "20-25", "25-25")),
                                        mat_temp = col_factor(levels = c("20", "25")),
                                        acc_temp = col_factor(levels = c("20", "25")),
                                        treatment = col_factor(NULL),
                                        exposed = col_factor(NULL),
                                        infected = col_factor(NULL)))

pheno_data <- pheno_data0 %>%
  filter(!is.na(death_date)) %>%
  mutate(status = 1)

glimpse(pheno_data)
```

##Load data: data for intrinsic growth rate calculations
```{r}
#data for calculating intrinsic growth rate
offspring_cum <- read_csv(file = "../data/full_offspring_data.csv",
                          col_names = TRUE,
                          col_types = list(host = col_factor(NULL),
                                           pathogen = col_factor(NULL),
                                           mat_temp = col_factor(NULL),
                                           acc_temp = col_factor(NULL),
                                           exposed = col_factor(NULL),
                                           birth_date = col_date("%Y-%m-%d"),
                                           inf_date = col_date("%Y-%m-%d")))

glimpse(offspring_cum)
```

##MCMC settings
```{r}
##### MCMC Settings
ni <- 75000
nb <- 30000
nt <- 16
nc <- 3
```

#####
##Death rate (little d) of uninfected hosts - observed all hosts die so no censor date needed
#####

#Subset data
```{r}
d_data_ctrl2020 <- subset(pheno_data, treatment == "CTRL-20-20")
d_data_ctrl2025 <- subset(pheno_data, treatment == "CTRL-20-25")
d_data_ctrl2520 <- subset(pheno_data, treatment == "CTRL-25-20")
d_data_ctrl2525 <- subset(pheno_data, treatment == "CTRL-25-25")
```


#Model
```{r}
sink("../JAGS_models/dmodelsingletempnocens.txt")
cat("
    model{
    
    ## Priors
    d ~ dunif(0, 1)
    
    ## Likelihood
    for(i in 1:N.obs.d){
    death[i] ~ dexp(d)
    }
    
    } # close model
    ",fill=T)
sink()

```

#Parameters to Estimate
```{r}
parameters <- c("d")
```

#Inits Function
```{r}
inits<-function(){list(
  d = 0.01)}
```


2020 treatments
```{r}
#Organize Data for JAGS - store data and run model for one temp at a time
death <- d_data_ctrl2020$lifespan_days 

#Length
N.obs.d <- length(death) 

#Bundle Data
jag.data<-list(
  death = death,
  N.obs.d = N.obs.d) 

#Run JAGS 
d_2020_out<-jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/dmodelsingletempnocens.txt",
                n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

2025 treatments
```{r}
death <- d_data_ctrl2025$lifespan_days

#Length
N.obs.d <- length(death) 

#Bundle Data
jag.data<-list(
  death = death,
  N.obs.d = N.obs.d) 


d_2025_out<-jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/dmodelsingletempnocens.txt",
                n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())


```

2520 treatments
```{r}
death <- d_data_ctrl2520$lifespan_days

#Length
N.obs.d <- length(death) 

#Bundle Data
jag.data<-list(
  death = death,
  N.obs.d = N.obs.d) 

d_2520_out<-jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/dmodelsingletempnocens.txt",
                n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

2525 treatments
```{r}
death <- d_data_ctrl2525$lifespan_days

#Length
N.obs.d <- length(death) 

#Bundle Data
jag.data<-list(
  death = death,
  N.obs.d = N.obs.d) 


d_2525_out<-jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/dmodelsingletempnocens.txt",
                n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

#Summary of output
```{r}
d_2020_out$BUGSoutput$summary
d_2025_out$BUGSoutput$summary
d_2520_out$BUGSoutput$summary
d_2525_out$BUGSoutput$summary
```

#Check diagnostics
```{r}
mcmcplot(d_2020_out)
mcmcplot(d_2025_out)
mcmcplot(d_2520_out)
mcmcplot(d_2525_out)
```

#Save summary data for each treatment
```{r}
JAGS_little_d_summary <- data.frame(temp_treat = c("20-20", "20-25", "25-20", "25-25"),
                                    median = c(median(d_2020_out$BUGSoutput$sims.list$d), median(d_2025_out$BUGSoutput$sims.list$d), median(d_2520_out$BUGSoutput$sims.list$d), median(d_2525_out$BUGSoutput$sims.list$d)), 
                                    mean = c(d_2020_out$BUGSoutput$summary[1, 1], d_2025_out$BUGSoutput$summary[1, 1], d_2520_out$BUGSoutput$summary[1, 1], d_2525_out$BUGSoutput$summary[1, 1]),
                                    lower = c(d_2020_out$BUGSoutput$summary[1, 3], d_2025_out$BUGSoutput$summary[1, 3], d_2520_out$BUGSoutput$summary[1, 3], d_2525_out$BUGSoutput$summary[1, 3]),
                                    upper = c(d_2020_out$BUGSoutput$summary[1, 7], d_2025_out$BUGSoutput$summary[1, 7], d_2520_out$BUGSoutput$summary[1, 7], d_2525_out$BUGSoutput$summary[1, 7]))

JAGS_little_d_summary
```

#Save fits for later calculations
```{r}
JAGS_little_d_2020 <- data.frame(little_d = d_2020_out$BUGSoutput$sims.list$d)
JAGS_little_d_2025 <- data.frame(little_d = d_2025_out$BUGSoutput$sims.list$d)
JAGS_little_d_2520 <- data.frame(little_d = d_2520_out$BUGSoutput$sims.list$d)
JAGS_little_d_2525 <- data.frame(little_d = d_2525_out$BUGSoutput$sims.list$d)
```




#####
##Intrinsic rate of growth
#####

First calculate little r for each individual

#Function for calculating intrinsic growth rate per individual for nested data
```{r}
# First write a contained function that works on a test df
# Coudl be much better, to allow for varying column names. But good enough.
fnc_el_uni <- function(df){
  
  r_range <- c(0, 5) 
  x_test <- c(df$host_age) 
  L_test <- c(df$alive_dead_status) 
  m_test <- c(df$offspring) 
  
  eulerlotka <- function(r) sum(L_test * m_test * exp(-r * x_test)) - 1 
  res <- uniroot(f = eulerlotka, interval = r_range, tol = 1e-8) 
  solution <- (res$root)
  
  solution
  
}
```

#subset and nest data
```{r}
ctrl_2020_nest <- offspring_cum %>% filter(pathogen == "CTRL", temp_treat == "20-20") %>%
  mutate(alive_dead_status = as.numeric(case_when(alive_dead == "ALIVE" ~ 1, alive_dead == "DEAD" ~ 0))) %>%
  select(indv_id, offspring, host_age, alive_dead_status) %>%
  group_by(indv_id) %>%
  nest()

ctrl_2025_nest <- offspring_cum %>% filter(pathogen == "CTRL", temp_treat == "20-25") %>%
  mutate(alive_dead_status = as.numeric(case_when(alive_dead == "ALIVE" ~ 1, alive_dead == "DEAD" ~ 0))) %>%
  select(indv_id, offspring, host_age, alive_dead_status) %>%
  group_by(indv_id) %>%
  nest()

ctrl_2520_nest <- offspring_cum %>% filter(pathogen == "CTRL", temp_treat == "25-20") %>%
  mutate(alive_dead_status = as.numeric(case_when(alive_dead == "ALIVE" ~ 1, alive_dead == "DEAD" ~ 0))) %>%
  select(indv_id, offspring, host_age, alive_dead_status) %>%
  group_by(indv_id) %>%
  nest()

ctrl_2525_nest <- offspring_cum %>% filter(pathogen == "CTRL", temp_treat == "25-25") %>%
  mutate(alive_dead_status = as.numeric(case_when(alive_dead == "ALIVE" ~ 1, alive_dead == "DEAD" ~ 0))) %>%
  select(indv_id, offspring, host_age, alive_dead_status) %>%
  group_by(indv_id) %>%
  nest()
```

#little r calcs
```{r}
little_r_2020 <- ctrl_2020_nest %>%
  mutate(little_r = map_dbl(data, fnc_el_uni)) %>%
  unnest(cols = little_r) %>%
  select(- data)

mean(little_r_2020$little_r)
```

```{r}
little_r_2025 <- ctrl_2025_nest %>%
  mutate(little_r = map_dbl(data, fnc_el_uni)) %>%
  unnest(cols = little_r) %>%
  select(- data)

mean(little_r_2025$little_r)
```

```{r}
little_r_2520 <- ctrl_2520_nest %>%
  mutate(little_r = map_dbl(data, fnc_el_uni)) %>%
  unnest(cols = little_r) %>%
  select(- data)

mean(little_r_2520$little_r)
```

```{r}
little_r_2525 <- ctrl_2525_nest %>%
  mutate(little_r = map_dbl(data, fnc_el_uni)) %>%
  unnest(cols = little_r) %>%
  select(- data)

mean(little_r_2525$little_r)
```

#jags
#General model for a normally distributed trait (for spores and r)
```{r}
sink("../JAGS_models/traitmumodelsingletemp.txt")
cat("
    model{
    
    ## Priors
    trait.mu ~ dnorm(0, 1/1000000000000)
    trait.sigma ~ dunif(0, 1000)
    trait.tau <- 1 / (trait.sigma * trait.sigma)
    
    ## Likelihood
    for(i in 1:N.obs.trait){
    trait[i] ~ dnorm(trait.mu, trait.tau)
    }
    
    } # close model
    ",fill=T)
sink()
```

#Parameters to Estimate
```{r}
parameters <- c("trait.mu", "trait.sigma")
```

# Inits Function
```{r}
inits<-function(){list(
  trait.mu = runif(1, 0, 1),
  trait.sigma = rlnorm(1))}
```

2020
```{r}
trait <- little_r_2020$little_r
N.obs.trait <- length(trait)
jag.data<-list(
  trait = trait,
  N.obs.trait = N.obs.trait)

r_2020_out <- jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/traitmumodelsingletemp.txt",
                 n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())

```

2025
```{r}
trait <- little_r_2025$little_r
N.obs.trait <- length(trait)
jag.data<-list(
  trait = trait,
  N.obs.trait = N.obs.trait)

r_2025_out <- jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/traitmumodelsingletemp.txt",
                 n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())

```

2520
```{r}
trait <- little_r_2520$little_r
N.obs.trait <- length(trait)
jag.data<-list(
  trait = trait,
  N.obs.trait = N.obs.trait)

r_2520_out <- jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/traitmumodelsingletemp.txt",
                 n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

2525
```{r}
trait <- little_r_2525$little_r
N.obs.trait <- length(trait)
jag.data<-list(
  trait = trait,
  N.obs.trait = N.obs.trait)

r_2525_out <- jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/traitmumodelsingletemp.txt",
                 n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

#Summary
```{r}
r_2020_out$BUGSoutput$summary
r_2025_out$BUGSoutput$summary
r_2520_out$BUGSoutput$summary
r_2525_out$BUGSoutput$summary
```

#Diagnostics
```{r}
mcmcplot(r_2020_out)
mcmcplot(r_2025_out)
mcmcplot(r_2520_out)
mcmcplot(r_2525_out)
```

#Save summary for plotting etc
```{r}
JAGS_little_r_summary <- data.frame(temp_treat = c("20-20", "20-25", "25-20", "25-25"),
                                    median = c(median(r_2020_out$BUGSoutput$sims.list$trait.mu), median(r_2025_out$BUGSoutput$sims.list$trait.mu), median(r_2520_out$BUGSoutput$sims.list$trait.mu), median(r_2525_out$BUGSoutput$sims.list$trait.mu)),
                                    mean = c(r_2020_out$BUGSoutput$summary[2, 1], r_2025_out$BUGSoutput$summary[2, 1], r_2520_out$BUGSoutput$summary[2, 1], r_2525_out$BUGSoutput$summary[2, 1]),
                                    lower = c(r_2020_out$BUGSoutput$summary[2, 3], r_2025_out$BUGSoutput$summary[2, 3], r_2520_out$BUGSoutput$summary[2, 3], r_2525_out$BUGSoutput$summary[2, 3]),
                                    upper = c(r_2020_out$BUGSoutput$summary[2, 7], r_2025_out$BUGSoutput$summary[2, 7], r_2520_out$BUGSoutput$summary[2, 7], r_2525_out$BUGSoutput$summary[2, 7]))

JAGS_little_r_summary
```

#Save JAGS output for later R0 trait calcs
```{r}
JAGS_little_r_2020 <- data.frame(little_r = r_2020_out$BUGSoutput$sims.list$trait.mu)
JAGS_little_r_2025 <- data.frame(little_r = r_2025_out$BUGSoutput$sims.list$trait.mu)
JAGS_little_r_2520 <- data.frame(little_r = r_2520_out$BUGSoutput$sims.list$trait.mu)
JAGS_little_r_2525 <- data.frame(little_r = r_2525_out$BUGSoutput$sims.list$trait.mu)
```





#####
##Transmission rate (beta)
#####

#transmission data
#wrangle data to get numbers of infected, uninfected and total for beta calcs

#total number for each treatment
```{r}
total_n <- pheno_data %>%
  filter(!is.na(infected)) %>%
  group_by(treatment) %>%
  summarise(Total = n())
```

#total infected for each treatment 
```{r}
inf_n <- pheno_data %>%
  filter(!is.na(infected)) %>%
  filter(infected != "N") %>%
  filter(infected != "P") %>%
  group_by(treatment) %>%
  summarise(Infe = n())
```

#merging and calculating number of uninfected animals in each treatment
```{r}
inf_prev_data <- total_n %>% left_join(inf_n, by = "treatment") %>%
  mutate(Uninf = Total - Infe)
```

#Changing C20-20-20 so it isnt 100% infection rate - making one individual uninfected
```{r}
inf_prev_data[3,4] <- 1
inf_prev_data[3,3] <- 16
inf_prev_data 
```

#subset transmission rate data by treatment
```{r}
beta_data_C12020 <- subset(inf_prev_data, treatment == "C1-20-20")
beta_data_C12025 <- subset(inf_prev_data, treatment == "C1-20-25")
beta_data_C12520 <- subset(inf_prev_data, treatment == "C1-25-20")
beta_data_C12525 <- subset(inf_prev_data, treatment == "C1-25-25")
beta_data_C142020 <- subset(inf_prev_data, treatment == "C14-20-20")
beta_data_C142025 <- subset(inf_prev_data, treatment == "C14-20-25")
beta_data_C142520 <- subset(inf_prev_data, treatment == "C14-25-20")
beta_data_C142525 <- subset(inf_prev_data, treatment == "C14-25-25")
beta_data_C202020 <- subset(inf_prev_data, treatment == "C20-20-20")
beta_data_C202025 <- subset(inf_prev_data, treatment == "C20-20-25")
beta_data_C202520 <- subset(inf_prev_data, treatment == "C20-25-20")
beta_data_C202525 <- subset(inf_prev_data, treatment == "C20-25-25")
```


# Model
```{r}
sink("../JAGS_models/betamodelsingletemp.txt")
cat("
    model{
    
    ## Priors
    beta ~ dnorm(0, 0.000001)
    
    ## Likelihood
    for(i in 1:N.obs){
    Num.uninf[i] ~ dbin(prob[i], Num[i])
    prob[i] <- exp(-beta * 2000 * 20 * 3 * 0.00001) 
    # 120000 total spore days = 2000 sp/ml x 20 ml x 3 day exposure
    # the beta parameter value is transformed by x 10^-5 to make fitting easier - need to transform output later
    }
    
    } # close model
    ",fill=T)
sink()
```

# Inits Function
```{r}
inits<-function(){list(
  beta = runif(1, 0, 1))}
```


# Parameters to Estimate
```{r}
parameters <- c("beta")
```


# Organize Data for JAGS - store data and run model for one temp at a time
```{r}
#C1
Num <- beta_data_C12020$Total
Num.uninf <- beta_data_C12020$Uninf
Num <- beta_data_C12025$Total
Num.uninf <- beta_data_C12025$Uninf
Num <- beta_data_C12520$Total
Num.uninf <- beta_data_C12520$Uninf
Num <- beta_data_C12525$Total
Num.uninf <- beta_data_C12525$Uninf
#C14
Num <- beta_data_C142020$Total
Num.uninf <- beta_data_C142020$Uninf
Num <- beta_data_C142025$Total
Num.uninf <- beta_data_C142025$Uninf
Num <- beta_data_C142520$Total
Num.uninf <- beta_data_C142520$Uninf
Num <- beta_data_C142525$Total
Num.uninf <- beta_data_C142525$Uninf
#C20
Num <- beta_data_C202020$Total
Num.uninf <- beta_data_C202020$Uninf
Num <- beta_data_C202025$Total
Num.uninf <- beta_data_C202025$Uninf
Num <- beta_data_C202520$Total
Num.uninf <- beta_data_C202520$Uninf
Num <- beta_data_C202525$Total
Num.uninf <- beta_data_C202525$Uninf
```

#length
```{r}
N.obs <- length(Num.uninf)
```

# Bundle Data
```{r}
jag.data<-list(
  Num = Num,
  Num.uninf = Num.uninf,
  N.obs = N.obs)
```

#Run JAGS one treatment at a time
```{r}
beta_C202525_out<-jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/betamodelsingletemp.txt",
                  n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

```{r}
beta_C12020_out$BUGSoutput$summary
beta_C12025_out$BUGSoutput$summary
beta_C12520_out$BUGSoutput$summary
beta_C12525_out$BUGSoutput$summary

beta_C142020_out$BUGSoutput$summary
beta_C142025_out$BUGSoutput$summary
beta_C142520_out$BUGSoutput$summary
beta_C142525_out$BUGSoutput$summary

beta_C202020_out$BUGSoutput$summary
beta_C202025_out$BUGSoutput$summary
beta_C202520_out$BUGSoutput$summary
beta_C202525_out$BUGSoutput$summary
```

```{r}
mcmcplot(beta_C12020_out)
mcmcplot(beta_C12025_out)
mcmcplot(beta_C12520_out)
mcmcplot(beta_C12525_out)

mcmcplot(beta_C142020_out)
mcmcplot(beta_C142025_out)
mcmcplot(beta_C142520_out)
mcmcplot(beta_C142525_out)

mcmcplot(beta_C202020_out)
mcmcplot(beta_C202025_out)
mcmcplot(beta_C202520_out)
mcmcplot(beta_C202525_out)
```

DONT FORGET I NEED TO CORRECT THE BETA BY "/(1/0.02)*0.00001" in both the summary and the raw posteriors

#saving summary of beta in data frame
```{r}
JAGS_beta_summary <- data.frame(temp_treat = c("20-20", "20-25", "25-20", "25-25", "20-20", "20-25", "25-20", "25-25", "20-20", "20-25", "25-20", "25-25"), 
                                treatment = c("C1-20-20", "C1-20-25", "C1-25-20", "C1-25-25", "C14-20-20", "C14-20-25", "C14-25-20", "C14-25-25", "C20-20-20", "C20-20-25", "C20-25-20", "C20-25-25"),
                                median = c(median(beta_C12020_out$BUGSoutput$sims.list$beta), median(beta_C12025_out$BUGSoutput$sims.list$beta), median(beta_C12520_out$BUGSoutput$sims.list$beta), median(beta_C12525_out$BUGSoutput$sims.list$beta), median(beta_C142020_out$BUGSoutput$sims.list$beta), median(beta_C142025_out$BUGSoutput$sims.list$beta), median(beta_C142520_out$BUGSoutput$sims.list$beta), median(beta_C142525_out$BUGSoutput$sims.list$beta), median(beta_C202020_out$BUGSoutput$sims.list$beta), median(beta_C202025_out$BUGSoutput$sims.list$beta), median(beta_C202520_out$BUGSoutput$sims.list$beta), median(beta_C202525_out$BUGSoutput$sims.list$beta)),
                                mean = c(beta_C12020_out$BUGSoutput$summary[1, 1], beta_C12025_out$BUGSoutput$summary[1, 1], beta_C12520_out$BUGSoutput$summary[1, 1], beta_C12525_out$BUGSoutput$summary[1, 1], beta_C142020_out$BUGSoutput$summary[1, 1], beta_C142025_out$BUGSoutput$summary[1, 1], beta_C142520_out$BUGSoutput$summary[1, 1], beta_C142525_out$BUGSoutput$summary[1, 1], beta_C202020_out$BUGSoutput$summary[1, 1], beta_C202025_out$BUGSoutput$summary[1, 1], beta_C202520_out$BUGSoutput$summary[1, 1], beta_C202525_out$BUGSoutput$summary[1, 1]),
                                lower = c(beta_C12020_out$BUGSoutput$summary[1, 3], beta_C12025_out$BUGSoutput$summary[1, 3], beta_C12520_out$BUGSoutput$summary[1, 3], beta_C12525_out$BUGSoutput$summary[1, 3], beta_C142020_out$BUGSoutput$summary[1, 3], beta_C142025_out$BUGSoutput$summary[1, 3], beta_C142520_out$BUGSoutput$summary[1, 3], beta_C142525_out$BUGSoutput$summary[1, 3], beta_C202020_out$BUGSoutput$summary[1, 3], beta_C202025_out$BUGSoutput$summary[1, 3], beta_C202520_out$BUGSoutput$summary[1, 3], beta_C202525_out$BUGSoutput$summary[1, 3]),
                                upper = c(beta_C12020_out$BUGSoutput$summary[1, 7], beta_C12025_out$BUGSoutput$summary[1, 7], beta_C12520_out$BUGSoutput$summary[1, 7], beta_C12525_out$BUGSoutput$summary[1, 7], beta_C142020_out$BUGSoutput$summary[1, 7], beta_C142025_out$BUGSoutput$summary[1, 7], beta_C142520_out$BUGSoutput$summary[1, 7], beta_C142525_out$BUGSoutput$summary[1, 7], beta_C202020_out$BUGSoutput$summary[1, 7], beta_C202025_out$BUGSoutput$summary[1, 7], beta_C202520_out$BUGSoutput$summary[1, 7], beta_C202525_out$BUGSoutput$summary[1, 7]))

JAGS_beta_summary #not corrected for density or 0.00001 yet
```

#Save beta posteriors for later calculations (untransformed so far)
```{r}
JAGS_beta_C12020 <- data.frame(beta = beta_C12020_out$BUGSoutput$sims.list$beta)
JAGS_beta_C12025 <- data.frame(beta = beta_C12025_out$BUGSoutput$sims.list$beta)
JAGS_beta_C12520 <- data.frame(beta = beta_C12520_out$BUGSoutput$sims.list$beta)
JAGS_beta_C12525 <- data.frame(beta = beta_C12525_out$BUGSoutput$sims.list$beta)

JAGS_beta_C142020 <- data.frame(beta = beta_C142020_out$BUGSoutput$sims.list$beta)
JAGS_beta_C142025 <- data.frame(beta = beta_C142025_out$BUGSoutput$sims.list$beta)
JAGS_beta_C142520 <- data.frame(beta = beta_C142520_out$BUGSoutput$sims.list$beta)
JAGS_beta_C142525 <- data.frame(beta = beta_C142525_out$BUGSoutput$sims.list$beta)

JAGS_beta_C202020 <- data.frame(beta = beta_C202020_out$BUGSoutput$sims.list$beta)
JAGS_beta_C202025 <- data.frame(beta = beta_C202025_out$BUGSoutput$sims.list$beta)
JAGS_beta_C202520 <- data.frame(beta = beta_C202520_out$BUGSoutput$sims.list$beta)
JAGS_beta_C202525 <- data.frame(beta = beta_C202525_out$BUGSoutput$sims.list$beta)
```







#####
###Spore yields
#####

Going to run on using spores divided by 100000 (one hundred thousand), need to multiply out to raw units later

# General model for a normally distributed trait (for sigma and r)
```{r}
sink("../JAGS_models/traitmumodelsingletemp.txt")
cat("
    model{
    
    ## Priors
    trait.mu ~ dnorm(0, 1/1000000000000)
    trait.sigma ~ dunif(0, 1000)
    trait.tau <- 1 / (trait.sigma * trait.sigma)
    
    ## Likelihood
    for(i in 1:N.obs.trait){
    trait[i] ~ dnorm(trait.mu, trait.tau)
    }
    
    } # close model
    ",fill=T)
sink()
```

# Parameters to Estimate
```{r}
parameters <- c("trait.mu", "trait.sigma")
```

#Inits Function
```{r}
inits<-function(){list(
  trait.mu = runif(1, 0, 30),
  trait.sigma = rlnorm(1))}
```

#Subset Data by Treatment
```{r}
spores_C12020 <- pheno_data %>% filter(treatment == "C1-20-20", infected == "M")
spores_C12025 <- pheno_data %>% filter(treatment == "C1-20-25", infected == "M")
spores_C12520 <- pheno_data %>% filter(treatment == "C1-25-20", infected == "M")
spores_C12525 <- pheno_data %>% filter(treatment == "C1-25-25", infected == "M")

spores_C142020 <- pheno_data %>% filter(treatment == "C14-20-20", infected == "M")
spores_C142025 <- pheno_data %>% filter(treatment == "C14-20-25", infected == "M")
spores_C142520 <- pheno_data %>% filter(treatment == "C14-25-20", infected == "M")
spores_C142525 <- pheno_data %>% filter(treatment == "C14-25-25", infected == "M")

spores_C202020 <- pheno_data %>% filter(treatment == "C20-20-20", infected == "M")
spores_C202025 <- pheno_data %>% filter(treatment == "C20-20-25", infected == "M")
spores_C202520 <- pheno_data %>% filter(treatment == "C20-25-20", infected == "M")
spores_C202525 <- pheno_data %>% filter(treatment == "C20-25-25", infected == "M")
```



#Organize Data for JAGS - store data and run model for one temp at a time
```{r}
trait <- spores_C12020$mature_spores/100000 
trait <- spores_C12025$mature_spores/100000 
trait <- spores_C12520$mature_spores/100000 
trait <- spores_C12525$mature_spores/100000 

trait <- spores_C142020$mature_spores/100000 
trait <- spores_C142025$mature_spores/100000 
trait <- spores_C142520$mature_spores/100000 
trait <- c(spores_C142525$mature_spores/100000, spores_C142525$mature_spores/100000) 

trait <- spores_C202020$mature_spores/100000 
trait <- c(spores_C202025$mature_spores/100000, spores_C202025$mature_spores/100000) 
trait <- spores_C202520$mature_spores/100000 
trait <- spores_C202525$mature_spores/100000 
```

#Length
```{r}
N.obs.trait <- length(trait)
```

#Bundle Data
```{r}
jag.data<-list(
  trait = trait,
  N.obs.trait = N.obs.trait)
```

#Run JAGS - change name each time you run the model
```{r}
spores_C202525_out <- jags(data=jag.data, inits=inits, parameters.to.save=parameters, model.file="../JAGS_models/traitmumodelsingletemp.txt",
                n.thin=nt, n.chains=nc, n.burnin=nb, n.iter=ni, DIC=T, working.directory=getwd())
```

# View parameters
```{r}
spores_C12020_out$BUGSoutput$summary
spores_C12025_out$BUGSoutput$summary
spores_C12520_out$BUGSoutput$summary
spores_C12525_out$BUGSoutput$summary

spores_C142020_out$BUGSoutput$summary
spores_C142025_out$BUGSoutput$summary
spores_C142520_out$BUGSoutput$summary
spores_C142525_out$BUGSoutput$summary #problem

spores_C202020_out$BUGSoutput$summary
spores_C202025_out$BUGSoutput$summary #problem
spores_C202520_out$BUGSoutput$summary
spores_C202525_out$BUGSoutput$summary

```

#diagnostics
```{r}
mcmcplot(spores_C12020_out)
mcmcplot(spores_C12025_out)
mcmcplot(spores_C12520_out)
mcmcplot(spores_C12525_out)

mcmcplot(spores_C142020_out)
mcmcplot(spores_C142025_out)
mcmcplot(spores_C142520_out)
mcmcplot(spores_C142525_out) #problem (only 5 infected)

mcmcplot(spores_C202020_out)
mcmcplot(spores_C202025_out) #problem (only 4 infected)
mcmcplot(spores_C202520_out)
mcmcplot(spores_C202525_out)
```

#Summary table of spore output (not corrected for transformation) - means might be more informative here than medians. In two cases the medians are way too high (C1-25-25 & C14-25-25)
```{r}
JAGS_spores_summary <- data.frame(treatment = c("C1-20-20", "C1-20-25", "C1-25-20", "C1-25-25", "C14-20-20", "C14-20-25", "C14-25-20", "C14-25-25", "C20-20-20", "C20-20-25", "C20-25-20", "C20-25-25"), 
                                  median = c(median(spores_C12020_out$BUGSoutput$sims.list$trait.mu), median(spores_C12025_out$BUGSoutput$sims.list$trait.mu), median(spores_C12520_out$BUGSoutput$sims.list$trait.mu), median(spores_C12025_out$BUGSoutput$sims.list$trait.mu), median(spores_C142020_out$BUGSoutput$sims.list$trait.mu), median(spores_C142025_out$BUGSoutput$sims.list$trait.mu), median(spores_C142520_out$BUGSoutput$sims.list$trait.mu), median(spores_C142025_out$BUGSoutput$sims.list$trait.mu), median(spores_C202020_out$BUGSoutput$sims.list$trait.mu), median(spores_C202025_out$BUGSoutput$sims.list$trait.mu), median(spores_C202520_out$BUGSoutput$sims.list$trait.mu), median(spores_C202025_out$BUGSoutput$sims.list$trait.mu)),
                                  mean = c(spores_C12020_out$BUGSoutput$summary[2, 1], spores_C12025_out$BUGSoutput$summary[2, 1], spores_C12520_out$BUGSoutput$summary[2, 1], spores_C12525_out$BUGSoutput$summary[2, 1], spores_C142020_out$BUGSoutput$summary[2, 1], spores_C142025_out$BUGSoutput$summary[2, 1], spores_C142520_out$BUGSoutput$summary[2, 1], spores_C142525_out$BUGSoutput$summary[2, 1], spores_C202020_out$BUGSoutput$summary[2, 1], spores_C202025_out$BUGSoutput$summary[2, 1], spores_C202520_out$BUGSoutput$summary[2, 1], spores_C202525_out$BUGSoutput$summary[2, 1]),
                                  lower = c(spores_C12020_out$BUGSoutput$summary[2, 3], spores_C12025_out$BUGSoutput$summary[2, 3], spores_C12520_out$BUGSoutput$summary[2, 3], spores_C12525_out$BUGSoutput$summary[2, 3], spores_C142020_out$BUGSoutput$summary[2, 3], spores_C142025_out$BUGSoutput$summary[2, 3], spores_C142520_out$BUGSoutput$summary[2, 3], spores_C142525_out$BUGSoutput$summary[2, 3], spores_C202020_out$BUGSoutput$summary[2, 3], spores_C202025_out$BUGSoutput$summary[2, 3], spores_C202520_out$BUGSoutput$summary[2, 3], spores_C202525_out$BUGSoutput$summary[2, 3]),
                                  upper = c(spores_C12020_out$BUGSoutput$summary[2, 7], spores_C12025_out$BUGSoutput$summary[2, 7], spores_C12520_out$BUGSoutput$summary[2, 7], spores_C12525_out$BUGSoutput$summary[2, 7], spores_C142020_out$BUGSoutput$summary[2, 7], spores_C142025_out$BUGSoutput$summary[2, 7], spores_C142520_out$BUGSoutput$summary[2, 7], spores_C142525_out$BUGSoutput$summary[2, 7], spores_C202020_out$BUGSoutput$summary[2, 7], spores_C202025_out$BUGSoutput$summary[2, 7], spores_C202520_out$BUGSoutput$summary[2, 7], spores_C202525_out$BUGSoutput$summary[2, 7]))

JAGS_spores_summary
```

#Saving raw posteriors for later derived parameter calculations
```{r}
JAGS_spores_C12020 <- data.frame(spore_load = spores_C12020_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C12025 <- data.frame(spore_load = spores_C12025_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C12520 <- data.frame(spore_load = spores_C12520_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C12525 <- data.frame(spore_load = spores_C12525_out$BUGSoutput$sims.list$trait.mu)

JAGS_spores_C142020 <- data.frame(spore_load = spores_C142020_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C142025 <- data.frame(spore_load = spores_C142025_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C142520 <- data.frame(spore_load = spores_C142520_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C142525 <- data.frame(spore_load = spores_C142525_out$BUGSoutput$sims.list$trait.mu)

JAGS_spores_C202020 <- data.frame(spore_load = spores_C202020_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C202025 <- data.frame(spore_load = spores_C202025_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C202520 <- data.frame(spore_load = spores_C202520_out$BUGSoutput$sims.list$trait.mu)
JAGS_spores_C202525 <- data.frame(spore_load = spores_C202525_out$BUGSoutput$sims.list$trait.mu)
```




###############
####Merging posteriors and calculating derived parameters
##############

##First combine posteriors of death rate (little d) and intrinsic rate of increase (little r) and calculate birth rate (little b), where b = r + d

#Merge posteriors for little r and little d and calculate little b
```{r}
JAGS_birth_death_2020 <- data.frame(temp_treat = "20-20", JAGS_little_r_2020, JAGS_little_d_2020) %>%
  mutate(little_b = little_r + little_d)

JAGS_birth_death_2025 <- data.frame(temp_treat = "20-25", JAGS_little_r_2025, JAGS_little_d_2025) %>%
  mutate(little_b = little_r + little_d)

JAGS_birth_death_2520 <- data.frame(temp_treat = "25-20", JAGS_little_r_2520, JAGS_little_d_2520) %>%
  mutate(little_b = little_r + little_d)

JAGS_birth_death_2525 <- data.frame(temp_treat = "25-25", JAGS_little_r_2525, JAGS_little_d_2525) %>%
  mutate(little_b = little_r + little_d)
```

#Combine uninfected parameter data frames for later saving
```{r}
JAGS_birth_death_full <- bind_rows(JAGS_birth_death_2020, JAGS_birth_death_2025, JAGS_birth_death_2520, JAGS_birth_death_2525)

JAGS_birth_death_full
```


#For each pathogen genotype by temp treat combination, combining birth/death, beta & spore posteriors 
```{r}
JAGS_data_C12020 <- data.frame(treatment = "C1-20-20", JAGS_birth_death_2020, JAGS_beta_C12020, JAGS_spores_C12020)
JAGS_data_C12025 <- data.frame(treatment = "C1-20-25", JAGS_birth_death_2025, JAGS_beta_C12025, JAGS_spores_C12025)
JAGS_data_C12520 <- data.frame(treatment = "C1-25-20", JAGS_birth_death_2520, JAGS_beta_C12520, JAGS_spores_C12520)
JAGS_data_C12525 <- data.frame(treatment = "C1-25-25", JAGS_birth_death_2525, JAGS_beta_C12525, JAGS_spores_C12525)

JAGS_data_C142020 <- data.frame(treatment = "C14-20-20", JAGS_birth_death_2020, JAGS_beta_C142020, JAGS_spores_C142020)
JAGS_data_C142025 <- data.frame(treatment = "C14-20-25", JAGS_birth_death_2025, JAGS_beta_C142025, JAGS_spores_C142025)
JAGS_data_C142520 <- data.frame(treatment = "C14-25-20", JAGS_birth_death_2520, JAGS_beta_C142520, JAGS_spores_C142520)
JAGS_data_C142525 <- data.frame(treatment = "C14-25-25", JAGS_birth_death_2525, JAGS_beta_C142525, JAGS_spores_C142525)

JAGS_data_C202020 <- data.frame(treatment = "C20-20-20", JAGS_birth_death_2020, JAGS_beta_C202020, JAGS_spores_C202020)
JAGS_data_C202025 <- data.frame(treatment = "C20-20-25", JAGS_birth_death_2025, JAGS_beta_C202025, JAGS_spores_C202025)
JAGS_data_C202520 <- data.frame(treatment = "C20-25-20", JAGS_birth_death_2520, JAGS_beta_C202520, JAGS_spores_C202520)
JAGS_data_C202525 <- data.frame(treatment = "C20-25-25", JAGS_birth_death_2525, JAGS_beta_C202525, JAGS_spores_C202525)
```

#Bind all datasets (rows) together for each pathgogen by temp treatment
```{r}
JAGS_full_data0 <- bind_rows(JAGS_data_C12020, JAGS_data_C12025, JAGS_data_C12520, JAGS_data_C12525, JAGS_data_C142020, JAGS_data_C142025, JAGS_data_C142520, JAGS_data_C142525, JAGS_data_C202020, JAGS_data_C202025, JAGS_data_C202520, JAGS_data_C202525)

JAGS_full_data0$temp_treat <- as_factor(JAGS_full_data0$temp_treat)
JAGS_full_data0$treatment <- as_factor(JAGS_full_data0$treatment)

glimpse(JAGS_full_data0)
```

#correcting units for beta and spore load
#beta needs to be corrected for units (so it is per L) and also transformed by 0.00001 which we used for fitting. 
#spores are currently in per 100000 units, need to convert to raw units
#also adding constants to the data: carrying capacity (c = 0.01) & spore degradation rate (m = 0.9) - Clay (2019) Within-Host Priority Effects Systematically Alter Pathogen Coexistence have measures of Pasteuria degredation between 0.0032 – 0.47. Could try setting to ~ 0.5 instead of 0.9
#finally for each row calculate the R0 value. R0 in our model is: R0 = ((b-d)/(bc))*((spores*beta)/m)
```{r}
JAGS_full_data <- JAGS_full_data0 %>%
  mutate(beta = beta/(1/0.02)*0.00001) %>%
  mutate(spore_load = spore_load*100000) %>%
  mutate(c = 0.01) %>%
  mutate(m = 0.9) %>%
  mutate(R0 = ((little_b - little_d)/(little_b * c)) * ((spore_load * beta)/m))

JAGS_full_data
```

#Saving JAGS birth/death (control) posteriors
#Saving JAGS pathogen R0 posteriors
```{r}
write_csv(JAGS_birth_death_full, "../data/JAGS_birth_death_post_full.csv")

write_csv(JAGS_full_data, "../data/JAGS_R0_derived_post_full.csv")
```
