---
title: "Individual phenotypic trait analysis"
---

## Packages

```{r}
library(tidyverse)
library(broom)
library(nlme)
library(car)
library(MASS)
library(emmeans)
library(DHARMa)
```

# Notes

In the data used here "mat_temp" is the maternal acclimation treatment, and "acc_temp" is the focal acclimation treatment.  


## Heat tolerance (knockdown time)

data 

```{r}
heat_data <- read_csv("../data/final_heat_data.csv", 
                      col_names = TRUE,
                      col_types = list(host = col_factor("M"),
                                       pathogen = col_factor(levels = c("CTRL", "C1", "C14", "C20")),
                                       temp_treat = col_factor(levels = c("20-20", "20-25", "25-20", "25-25")),
                                       mat_temp = col_factor(levels = c("20", "25")),
                                       acc_temp = col_factor(levels = c("20", "25")),
                                       run = col_factor(levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")))) 

heat_data
```

Manually setting contrasts

```{r}
heat_data2 <- heat_data %>% 
  mutate(mat_acc_cont = mat_temp,
         acc_temp_cont = acc_temp,
         pathogen_cont = pathogen)

contrasts(heat_data2$mat_acc_cont) = contr.sum(2)
contrasts(heat_data2$acc_temp_cont) = contr.sum(2)
contrasts(heat_data2$pathogen_cont) = contr.sum(4)
```


Model

```{r echo=TRUE}
heat_mod <- lme(death_time ~ mat_acc_cont * acc_temp_cont * pathogen_cont, random = ~1|run, method = "REML", weights = varIdent(form = ~1|acc_temp), data = heat_data2)
```


```{r}
#plot(heat_mod)
#qqnorm(heat_mod$residuals)
#aa <- augment(heat_mod_3)
#ggplot(aa,aes(acc_temp,.resid))+
#    geom_boxplot()+coord_flip()
```

```{r}
summary(heat_mod)
```

```{r}
Anova(heat_mod, type = 3)
```



# Fitness traits

### data 

```{r}
pheno_data <- read_csv("../data/final_phenotype_data.csv",
                      col_names = TRUE,
                      col_types = list(host = col_factor(NULL),
                                       pathogen = col_factor(levels = c("CTRL", "C1", "C14", "C20")),
                                       temp_treat = col_factor(levels = c("20-20", "25-20", "20-25", "25-25")),
                                       mat_temp = col_factor(levels = c("20", "25")),
                                       acc_temp = col_factor(levels = c("20", "25")),
                                       treatment = col_factor(NULL),
                                       exposed = col_factor(NULL),
                                       infected = col_factor(NULL)))

glimpse(pheno_data)
```


## Host Lifespan 

Model 

```{r}
lifespan_lm <- lm(log(lifespan_days) ~ mat_temp * acc_temp * pathogen, contrasts = list(mat_temp = contr.Sum, acc_temp = contr.Sum, pathogen = contr.Sum), data = pheno_data)
```

```{r}
#plot(lifespan_lm)
```

Hypothesis test

```{r}
Anova(lifespan_lm, type = 3, white.adjust = TRUE)
```


## Host Lifetime Fecundity

```{r}
fecundity_lm <- lm(log(total_offspring) ~ mat_temp * acc_temp * pathogen, contrasts = list(mat_temp = contr.Sum, acc_temp = contr.Sum, pathogen = contr.Sum), data = pheno_data)
```


```{r}
Anova(fecundity_lm, type = 3, white.adjust = TRUE)
```


## Pathogen Spore Loads

Subsetting only hosts infected with mature transmission spores

```{r}
mature_spore_data <- pheno_data %>% filter(pathogen != "CTRL" & infected == "M")
```

```{r}
mature_spore_lm <- lm(mature_spores ~ mat_temp * acc_temp * pathogen, contrasts = list(mat_temp = contr.Sum, acc_temp = contr.Sum, pathogen = contr.Sum), data = mature_spore_data)
```

```{r}
#plot(mature_spore_lm)
```

```{r}
Anova(mature_spore_lm, type = 3, white.adjust = TRUE)
```


## Pathogen infection success

Coding individuals so they are yes or no for infections with mature spores
```{r}
infection_data <- pheno_data %>% filter(pathogen != "CTRL" & !is.na(infected)) %>%
  mutate(mature_spores_yn = as.numeric(case_when(infected == "M" ~ "1", TRUE ~ "0")))
```

For treatments with 100% infection, making one individual uninfected to allow for proper estimation of errors

C20-20-20- 100% for all spores and mature spores
```{r}
C20_2020_binom <- infection_data %>%
  filter(treatment == "C20-20-20") %>%
  mutate(mature_spores_yn = as.numeric(case_when(indv_id == "C20-20-20-34" ~ "0", TRUE ~ "1")))
```

removing old data and then joining above data
```{r}
infection_data_ammended <- infection_data %>% 
  filter(treatment != "C20-20-20") %>%
  bind_rows(C20_2020_binom)

infection_data_ammended
```


binomial glm for mature spores (Y/N)

```{r}
infect_rate_ms_glm <- glm(mature_spores_yn ~ mat_temp * acc_temp * pathogen, contrasts = list(mat_temp = contr.Sum, acc_temp = contr.Sum, pathogen = contr.Sum), family = binomial(link = "logit"), data = infection_data_ammended)
```

```{r}
glm_inf_sim <- simulateResiduals(fittedModel = infect_rate_ms_glm, plot = TRUE)
```


```{r}
Anova(infect_rate_ms_glm, type = 3)
```

Calculating mean (probability) and SE
```{r}
mean_inf_rate_ms <- as_tibble(emmeans(infect_rate_ms_glm, ~ mat_temp * acc_temp * pathogen, trans = "response")) %>%
  unite(treat_temp, acc_temp, mat_temp, sep = "-", remove = FALSE)

mean_inf_rate_ms
```


```{r}
#write_csv(mean_inf_rate_ms, "../data/infection_prob_data.csv")
```

